---
title: "CustomerChurn"
output: html_document
date: "2023-08-09"
---

AIRLINE CUSTOMER CHURN ANALYSIS
The following dataset was taken from kaggle. The objective of this exercise is 
1. Find out reasons for customer churn
2. Recommendations & Business insights to combat customer churn

Utilising the below cell block to import necessary library and packages

```{r setup, include=FALSE}
## Importing packages & dataset
library(dplyr)
library(stringr)
library(tidyverse)
library(skimr)
library(ggplot2)

flight_data<-read.csv("airline_passenger_satisfaction.csv")
print(head(flight_data))


```




```{r cars}
summary(flight_data)
#skim(flight_data)
```
```{r}
# Doing data-coversions

flight_data$Gender <- as.factor(flight_data$Gender)
flight_data$Customer.Type < -as.factor(flight_data$Customer.Type)

# or

# List of column names to convert to factors
columns_to_convert <- c("Type.of.Travel","Class","Satisfaction")

# Convert the specified columns to factor type
flight_data[columns_to_convert] <- sapply(flight_data[columns_to_convert], factor)

summary(flight_data)


```
```{r}
##Finding NAs in our columns

# Find NAs in each column
nas_in_columns <- sapply(flight_data, function(col) sum(is.na(col)))

# Print the count of NAs in each column
print(nas_in_columns)


```
```{r}
##Since only 393, we can go with imputation aswell, but in this case, i decided to remove the NAs
flight_data <- flight_data[!is.na(flight_data$Arrival.Delay),]

```


```{r}
# Finding duplicates - veryfing the ID column is unique
length(unique(flight_data$ID))


##checking for duplicates ignoring ID column
duplicates <- flight_data %>%
  select(-ID) %>%
  duplicated()

flight_data[duplicates,]

```

## EDA

Checking response column

```{r pressure, echo=FALSE}
# table(flight_data$Satisfaction)
tabyl(flight_data$Satisfaction)

ggplot(flight_data, aes(x = Satisfaction)) +
  geom_bar()

```
```{r}
# Doing Bivariate tables & graphs
## Gender satisfaction

k<- flight_data %>%
  group_by(Gender,Satisfaction) %>%
  summarise(count = n())
  
ggplot(k, aes(x=Gender,y = count,fill = Satisfaction)) + geom_bar(stat = 'identity') +  labs(title = "Stacked Bar Plot of Satisfaction by Gender",
       x = "Gender", y = "Count")
  
  
## Getting percentage among rows
k <- table(flight_data$Gender,flight_data$Satisfaction)
row_sum <- rowSums(k)
k/row_sum

## Plotting 
```
```{r}
## How satisfaction changes as per age. 
## First lets understand age distribution

flight_data$Age <- as.numeric(flight_data$Age)
ggplot(flight_data,aes(x = Age)) +  geom_histogram(binwidth = 10)

## Majority of Age between 20 to 60. Let us bin age and then plot a 100% stacked to understand satisfaction rates

flight_data$Age_Binned <- cut(flight_data$Age, breaks = seq(0, 100, by = 10), right = FALSE)

k1 <- flight_data %>%
  group_by(Age_Binned,Satisfaction) %>%
  summarise(count = n())

total_counts <- k1 %>%
  group_by(Age_Binned) %>%
  summarise(total_count = sum(count))

# Filter to keep only unique Age_Binned values
k1_with_total_unique <- k1_with_total %>%
  select(Age_Binned,Satisfaction,total_count) %>%
  distinct(Age_Binned,.keep_all = TRUE) 

colnames(k1_with_total_unique)[3] <- "count"

k1_with_total <- left_join(k1, total_counts, by = "Age_Binned")

ggplot(k1_with_total,aes(x=Age_Binned,y = count, fill= Satisfaction)) +geom_bar(stat='identity',position = 'Fill') + geom_text(data = k1_with_total_unique, aes(label = count),position = position_fill(vjust = 1.1)) +
scale_y_continuous(labels = scales::percent_format(scale = 1))



```

```{r}

  # For data splitting

# Assuming you have your data loaded into your_data
# and your target variable is named 'Class' (replace with your actual column name)


# Split data into training and testing sets
set.seed(123)  # for reproducibility
trainIndex <- createDataPartition(flight_data$Satisfaction, p = 0.7, list = FALSE)
train_data <- flight_data[trainIndex, ]
test_data <- flight_data[-trainIndex, ]

# Train the logistic regression
logreg_model <- glm(Satisfaction ~ .-ID-Age_Binned, data = train_data, family = "binomial")
summary(logreg_model)


```
```{r}
# Make predictions on the training data
predictions <- predict(logreg_model, type = "response",newdata = test_data)
# Convert predictions to binary classes
predicted_classes <- ifelse(predictions > 0.5, "Satisfied", "Not Satisfied")


# Evaluate the model's performance
conf_matrix <- table(predicted_classes, test_data$Satisfaction)
print(conf_matrix)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
print(paste("Accuracy:", accuracy))

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
