---
title: "CustomerChurn"
output: html_document
date: "2023-08-09"
---

AIRLINE CUSTOMER CHURN ANALYSIS

The following dataset was taken from kaggle. The objective of this exercise is 

1. Find out reasons for customer churn


```{r setup, include=FALSE}
## Importing packages & dataset
library(dplyr)
library(stringr)
library(tidyverse)
library(skimr)
library(ggplot2)
library(caret)

flight_data<-read.csv("airline_passenger_satisfaction.csv")
print(head(flight_data))


```




```{r cars}
# summary(flight_data)
skim(flight_data)
```
```{r}
# Doing data-coversions

flight_data$Gender <- as.factor(flight_data$Gender)

# or

# List of column names to convert to factors
columns_to_convert <- c("Type.of.Travel","Class","Satisfaction")

# Convert the specified columns to factor type
flight_data[columns_to_convert] <- sapply(flight_data[columns_to_convert], factor)

summary(flight_data)


```
```{r}
##Finding NAs in our columns

# Find NAs in each column
nas_in_columns <- sapply(flight_data, function(col) sum(is.na(col)))

# Print the count of NAs in each column
print(nas_in_columns)


```
```{r}
##Since only 393, we can go with imputation aswell, but in this case, i decided to remove the NAs
flight_data <- flight_data[!is.na(flight_data$Arrival.Delay),]

```


```{r}
# Finding duplicates - veryfing the ID column is unique
length(unique(flight_data$ID))


##checking for duplicates ignoring ID column
duplicates <- flight_data %>%
  select(-ID) %>%
  duplicated()

flight_data[duplicates,]

```

## EDA

Checking response column

```{r pressure, echo=FALSE}
table(flight_data$Satisfaction)
# tabyl(flight_data$Satisfaction)

ggplot(flight_data, aes(x = Satisfaction)) +
  geom_bar()

```
```{r}
# Doing Bivariate tables & graphs
## Gender satisfaction

k<- flight_data %>%
  group_by(Gender,Satisfaction) %>%
  summarise(count = n())
print(k)
  
ggplot(k, aes(x=Gender,y = count,fill = Satisfaction)) + geom_bar(stat = 'identity') +  labs(title = "Stacked Bar Plot of Satisfaction by Gender",
       x = "Gender", y = "Count")
  
  
## Getting percentage among rows
k <- table(flight_data$Gender,flight_data$Satisfaction)
row_sum <- rowSums(k)
k/row_sum

## Plotting 
```
```{r}
## How satisfaction changes as per age. 
## First lets understand age distribution

flight_data$Age <- as.numeric(flight_data$Age)

## Majority of Age between 20 to 60. Let us bin age and then plot a 100% stacked to understand satisfaction rates

flight_data$Age_Binned <- cut(flight_data$Age, breaks = seq(0, 100, by = 10), right = FALSE)

k1 <- flight_data %>%
  group_by(Age_Binned,Satisfaction) %>%
  summarise(count = n())

total_counts <- k1 %>%
  group_by(Age_Binned) %>%
  summarise(total_count = sum(count))

k1_with_total <- left_join(k1, total_counts, by = "Age_Binned")
print(k1_with_total)

# Filter to keep only unique Age_Binned values
k1_with_total_unique <- k1_with_total %>%
  select(Age_Binned,Satisfaction,total_count) %>%
  distinct(Age_Binned,.keep_all = TRUE) 

colnames(k1_with_total_unique)[3] <- "count"


ggplot(k1_with_total,aes(x=Age_Binned,y = count, fill= Satisfaction)) +geom_bar(stat='identity',position = 'Fill') + geom_text(data = k1_with_total_unique, aes(label = count),position = position_fill(vjust = 1.1)) +
scale_y_continuous(labels = scales::percent_format(scale = 1))



```

```{r}

# Convert response into 1 or 0
table(flight_data$Satisfaction)
flight_data$Satisfaction <- recode(flight_data$Satisfaction,"Neutral or Dissatisfied"=0,"Satisfied"=1)

# Split data into training and testing sets
set.seed(123)  # for reproducibility
trainIndex <- createDataPartition(flight_data$Satisfaction, p = 0.7, list = FALSE)
train_data <- flight_data[trainIndex, ]
test_data <- flight_data[-trainIndex, ]

# Train the logistic regression
logreg_model <- glm(Satisfaction ~ .-ID-Age_Binned, data = train_data, family = "binomial")
summary(logreg_model)


```
```{r}
# Make predictions on the training data
predictions <- predict(logreg_model, type = "response",newdata = test_data)
# Convert predictions to binary classes
predicted_classes <- ifelse(predictions > 0.5, 1, 0)


# Evaluate the model's performance
conf_matrix <- table(predicted_classes, test_data$Satisfaction)
print(conf_matrix)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
print(paste("Accuracy:", accuracy))

```

Reasons for customer churn:

1. Class Economy Plus & Economy customers are more likely to be detractor
2. Personal Type of travel customers are more likely to be detractors
3. If a customer's departure delay decreases, he is more likely to be promoter


```{r}
## Quality Checking the results with contigency tables
k1 <- table(flight_data$Class,flight_data$Satisfaction)
print(k1/rowSums(k1))

## 
k1 <- table(flight_data$Type.of.Travel,flight_data$Satisfaction)
print(k1/rowSums(k1))


flight_data$Departure_Delay_binned <- cut(flight_data$Departure.Delay, breaks = seq(0, 500, by = 50), right = FALSE)
k1 <- table(flight_data$Departure_Delay_binned,flight_data$Satisfaction)
k2<- k1[, "1"] / (k1[, "1"] + k1[, "0"])
print(cbind(k1, satisfaction_rate = k2))




```

